Effectiveness of community-based TURF-reserves in small-scale fisheries

### 2018-12-07 docs/manuscript.Rmd
Added latest pdf version to review by hand

In an effort to bridge this normative gap, Mexican Civil Society Organizations (CSOs) served as a link between fishers and government, and created a legal framework that solves this governance issue (*i.e.* Fish Refuges; \citet{nom}). Fish refuges can be implemented as temporal or partial reserves, which can protect one, some, or all resources within their boundaries. One of the ways in which fishing communities have takane advantage of this new tool is by implementing marine reserves within their TURFs. Our work focuses on some of these community-based TURF-reserves that occur in small-scale fisheries.

@@

The other two TURF-reserves are located in Maria Elena and Punta Herrero, in the Yucatan Peninsula (Fig \ref{fig:map}C). In contrast with Isla Nativdad, which hosts a well established fishing community, Maria Elena is a fishing camp --visited intermittently during the fishing season-- belonging to the *Cozumel* fishing cooperative; Punta Herrero is home to the *José María Azcorra* fishing cooperative, and similar to Isla Natividad hosts a local community. Their main fishery is the Caribbean spiny lobster (*Panulirus argus*), but they also target finfish in the off-season. Maria Elena and Punta Herrero established eight and four marine reserves in 2012 and 2013, respectively. These reserves have been legally recognized as Fishing Refuges since their creation \citep{dof_website_2012,dof_website_2013}.

>>>>

In an effort to bridge this normative gap, Mexican Civil Society Organizations (CSOs) served as a link between fishers and government, and created a legal framework that solves this governance issue (*i.e.* Fish Refuges; \citet{nom}). Fish refuges can be implemented as temporal or partial reserves, which can protect one, some, or all resources within their boundaries. One of the ways in which fishing communities have taken advantage of this new tool is by implementing temporal marine reserves within their TURFs with a defined expiration date (often 5 years). After these five years have passed, fishers have the opportunity to open the reserves to fishing or continue to have them. Our work focuses on some of these community-based TURF-reserves that occur in small-scale fisheries.

@@

The other two TURF-reserves are located in Maria Elena and Punta Herrero, in the Yucatan Peninsula (Fig \ref{fig:map}C). In contrast with Isla Nativdad, which hosts a well established fishing community, Maria Elena is a fishing camp --visited intermittently during the fishing season-- belonging to the *Cozumel* fishing cooperative; Punta Herrero is home to the *José María Azcorra* fishing cooperative, and similar to Isla Natividad hosts a local community. Their main fishery is the Caribbean spiny lobster (*Panulirus argus*), but they also target finfish in the off-season. Maria Elena and Punta Herrero established eight and four marine reserves in 2012 and 2013, respectively. These reserves have been legally recognized as Fishing Refuges since their original implementation \citep{dof_website_2012,dof_website_2013} and re-enactments \citep{dof_website_2017}.

### 2018-12-07 docs/manuscript.tex
Added latest pdf version to review by hand

Our approach to evaluate the temporal and spatial changes provides a
more robust measure of reserve effectiveness. For example, we capture
previously described patterns like the rapid increase observed for
lobster densities in Isla Natividad on the sixth year (\emph{i.e.} 2012;
Fig. \ref{fig:indicators}A), a year after the hypoxia events described
by \citet{micheli_2012-EU}, which caused mass mortality of sedentary
organisms such as abalone and sea urchins, but not lobster and finfish.
Yet, our empirical approach assumes control sites are a plausible
counterfactual for treated sites. This implies that treated sites would
have followed the same trend as control sites, had the reserves not been
implemented. Nonetheless, temporal trends for each site don't show any
significant increases (S1 Table, S2 Table, S3 Table), supporting our
findings of lack of change in the indicators used.

Our analyses of socioeconomic indicators has two limitations. First, we
only look at landings and revenues by landings for communities with and
without TURF-reserves. There are a number of other possible indicators
that could show a change due to the implementation of the reserve.
Notably, one often cited in the literature is additional benefits, such
as tourism \citep{viana_2017}. However, it is unlikely that the
evaluated communities will experience tourism benefits due to their
remoteness and the lack of proper infrastructure to sustain tourism.

A second limitation of our socioeconomic analysis is that we do not
observe effort data, which may mask the effect of the reserve. For
example, if catches remain relatively unchanged but fishing effort
decreased, that would imply a larger catch per unit effort. Even without
changes to revenues, a lower input of effort (\emph{e.g.} less fishing
hours, less trips, less fuel) would imply that the fishery is more
profitable. However, fishers are assambled into fishing cooperatives,
meaning that profits are shared amongst all members
\citep{mccay_2014,mccay_2017}. Therefore, the same revenue is still
shared by the same amount of fishers.

@@

Another key condition for effectiveness is reserve size
\citep{edgar_2014-UO}, and the lack of effectiveness can perhaps be
attributed to poor ecological coherence in reserve design (\emph{sensu}
\citet{rees_2018}). Previous research has shown that reserves in Isla
Natividad yield fishery benefits for the abalone fishery
\citep{rossetto_2015-V0}. Abalone are less mobile than lobsters, and
perhaps the reserves provide enough protection to these sedentary
invertebrates, but not lobsters. Design principles developed by
\citet{green_2017} for marine reserves in the Caribbean state that
reserves ``should be more than twice the size of the home range of
adults and juveniles'', and suggest that reserves seeking to protect
spiny lobsters should have at least 14 km across. Furthermore, fishers
may favor implementation of reserves that pose low fishing costs due to
their small size or location. Our analysis of economic data supports
this hypothesis, as neither landings nor revenues showed the expected
short-term costs associated to the first years of reserve implementation
\citep{ovando_2016-Wg}.

@@

While the evaluated reserves have failed to provide fishery benefits up
to now, there are a number of additional ecological, fisheries, and
social benefits. Marine reserves provide protection to a wider range of
species and vulnerable habitat. These sites can serve as an insurance
against uncertainty and errors in fisheries management, as well as mild
environmental shocks
\citep{micheli_2012-EU,deleo_2015,roberts_2017-J9,aalto}.
Self-regulation of fishing effort (\emph{i.e.} reduction in harvest) can
serve as a way to compensate for future declines associated to
environmental variation \citep{finkbeiner_2018}. Embarking in a marine
conservation project can bring the community together, which promotes
social cohesion and builds social capital \citep{fulton_2019}. Showing
commitment to marine conservation and sustainable fishing practices
allows fishers to have greater bargaining power and leverage over
fisheries management \citep{prezramrez_2012}. Furthermore, the lack of
effectiveness observed in these reserves should not be generalizable to
other reserves established under the same legal framework (\emph{i.e.}
Fish Refuges) in Mexico, and future research should aim at evaluating
other areas that have also been established as bottom-up processes but
without the presence of TURFs (\emph{e.g.} \citet{dof_websiteC_2012}),
or others established through a top-down process (\emph{i.e.}
\citet{dof_websiteU_2018}).

Community-based marine reserves in small-scale fisheries can be helpful
conservation and fishery management tools when appropriately
implemented. Lessons learned from these cases can guide implementation
of community-based marine reserves elsewhere. For the particular case of
the marine reserves that we evaluate, the possibility of expanding
reserves or merging existing polygons into larger areas should be
evaluated and proposed to the communities. Community-based marine
reserves might have more benefits that result from indirect effects of
the reserves, particularly providing resilience to shocks and management
errors, and promoting social cohesion, which should be taken into
account when evaluating the outcomes of TURF-reserves. Having full
community support surely represents an advantage, but it is important
that community-based TURF-reserves meet essential design principles such
as size and placement so as to maximize their effectiveness.

>>>>

Our approach to evaluate the temporal and spatial changes provides a
more robust measure of reserve effectiveness. For example, we capture
previously described patterns like the rapid increase observed for
lobster densities in Isla Natividad on the sixth year (\emph{i.e.} 2012;
Fig. \ref{fig:indicators}A), a year after the hypoxia events described
by \citet{micheli_2012-EU}, which caused mass mortality of sedentary
organisms such as abalone and sea urchins, but not lobster and finfish.
Yet, our empirical approach assumes control sites are a plausible
counterfactual for treated sites. This implies that treated sites would
have followed the same trend as control sites, had the reserves not been
implemented. Nonetheless, temporal trends for each site don't show any
significant increases (Figures S1-S4), supporting our findings of lack
of change in the indicators used.

Our analyses of socioeconomic indicators has two limitations. First, we
only look at landings and revenues by landings for communities with and
without TURF-reserves. There are a number of other possible indicators
that could show a change due to the implementation of the reserve.
Notably, one often cited in the literature is additional benefits, such
as tourism \citep{viana_2017}. However, it is unlikely that the
evaluated communities will experience tourism benefits due to their
remoteness and the lack of proper infrastructure to sustain tourism. A
second limitation of our socioeconomic analysis is that we do not
observe effort data, which may mask the effect of the reserve. For
example, if catches remain relatively unchanged but fishing effort
decreased, that would imply a larger catch per unit effort. Even without
changes to revenues, a lower input of effort (\emph{e.g.} less fishing
hours, less trips, less fuel) would imply that the fishery is more
profitable. However, fishers are assambled into fishing cooperatives,
meaning that profits are shared amongst all members
\citep{mccay_2014,mccay_2017}. Therefore, the same revenue is still
shared by the same amount of fishers.

@@

Another key condition for effectiveness is reserve size
\citep{edgar_2014-UO}, and the lack of effectiveness can perhaps be
attributed to poor ecological coherence in reserve design (\emph{sensu}
\citet{rees_2018}). Previous research has shown that reserves in Isla
Natividad yield fishery benefits for the abalone fishery
\citep{rossetto_2015-V0}. Abalone are less mobile than lobsters, and
perhaps the reserves provide enough protection to these sedentary
invertebrates, but not lobsters. Design principles developed by
\citet{green_2017} for marine reserves in the Caribbean state that
reserves ``should be more than twice the size of the home range of
adults and juveniles'', and suggest that reserves seeking to protect
spiny lobsters should have at least 14 km across. Furthermore, fishers
may favor implementation of reserves that pose low fishing costs due to
their small size or location. Our analysis of economic data supports
this hypothesis, as neither landings nor revenues showed the expected
short-term reductions associated to the first years of reserve
implementation \citep{ovando_2016-Wg}.

@@

While the evaluated reserves have failed to provide fishery benefits up
to now, there are a number of additional ecological, fisheries, and
social benefits. Marine reserves provide protection to a wider range of
species and vulnerable habitat. These sites can serve as an insurance
against uncertainty and errors in fisheries management, as well as mild
environmental shocks
\citep{micheli_2012-EU,deleo_2015,roberts_2017-J9,aalto}.
Self-regulation of fishing effort (\emph{i.e.} reduction in harvest) can
serve as a way to compensate for future declines associated to
environmental variation \citep{finkbeiner_2018}. Embarking in a marine
conservation project can bring the community together, which promotes
social cohesion and builds social capital \citep{fulton_2019}. Showing
commitment to marine conservation and sustainable fishing practices
allows fishers to have greater bargaining power and leverage over
fisheries management \citep{prezramrez_2012}. These additional benefits
might explain why communities show a positive perception about their
performance and continue to support their presence \citep{ayer_2018}.

In terms of the fisheries regulation in Mexico, our work only evaluates
Fish Refuges established within TURFs. Future research should aim at
evaluating other Fish Refuges that have also been established as
bottom-up processes but without the presence of TURFs (\emph{e.g.}
\citet{dof_websiteC_2012}), others established through top-down
processes (\emph{i.e.} \citet{dof_websiteU_2018}), as well as the
relationship between governance and effectiveness across this gradient
of approaches.

Community-based TURF-reserves in small-scale fisheries may be helpful
conservation and fishery management tools when appropriately implemented
\citep{gelcich_2015}. To implement them we must promote bottom-up design
and implementation processes like the ones in the evaluated reserves,
but without setting design principles aside. Lessons learned from these
cases can guide implementation of community-based TURF-reserves
elsewhere. For the particular case of the ones we evaluate, the
possibility of expanding reserves or merging existing polygons into
larger areas should be evaluated and proposed to the communities.
Community-based marine reserves might have more benefits that result
from indirect effects of the reserves, particularly providing resilience
to shocks and management errors, and promoting social cohesion, which
should be taken into account when evaluating the outcomes of
TURF-reserves. Having full community support surely represents an
advantage, but it is important that community-based TURF-reserves meet
essential design principles such as size and placement so as to maximize
their effectiveness.
